---
title: "midterm_practice"
author: "Kodiak Soled"
date: "10/17/2019"
output: github_document
always_allow_html: yes
---

```{r}
library(readxl)
library(tidyverse)
library(kableExtra)
library(viridis)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d

scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
wordcountaddin::text_stats("midterm_practice.Rmd")
```

# Problem 1

## Import/Clean Data

* Read in file
* `clean_names`
* `Replace` "eop_size" missing values with '0'
* Create appropriate character/factor categorical variables (`as.character`/`as.factor`)
* `Recode` "sex", "age_group", "eop_size", "eop_visibility_classification" and "fhp_category" to reflect definitions (e.g., '7'/'8' recoded as '6')
* `Recode` miscoded variables (causes problems with figures later on):
  * Participant #1153 changed from age_group '1' to '4'
  * Participant #686 changed eop_size = '14.6' to 'yes' for "eeop"
* Create "eeop" variable
* `Rename` variables during `recode`'ing 

```{r}
mtp_data = 
  read_excel("./data/p8105_mtp_data.xlsx", sheet = "this one", range = "A9:I1230") %>%
  janitor::clean_names() %>%
  mutate(
    eop_size_mm = replace(eop_size_mm, is.na(eop_size_mm) == TRUE, 0),
    sex = as.character(sex),
    fhp_category = as.factor(fhp_category), 
   eop_visibility_classification = as.factor(eop_visibility_classification),
    eop_size = as.factor(eop_size),
    age_group = as.factor(age_group),
    sex = recode(sex,
                 `1` = "male",
                 `0` = "female"),
    age_group = recode(age_group,
                       `7` = "6",
                       `8` = "6",
                       `1` = "4"),
    age_category = recode(age_group,
                       `2` = "18-30",
                       `3` = "31-40",
                       `4` = "41-50",
                       `5` = "51-60",
                       `6` = "61+"),
    age_category = forcats::fct_relevel(age_category, c("18-30", "31-40", "41-50", "51-60", "61+")),
    eeop = recode(eop_size,
                  `0` = "no",
                  `1` = "no",
                  `2` = "yes",
                  `3` = "yes",
                  `4` = "yes",
                  `5` = "yes", 
                  `14.6` = "yes"), 
   eop_category = recode(eop_size,
                         `0` = "0-5mm",
                         `1` = "5-10mm",
                         `2` = "10-15mm",
                         `3` = "15-20mm",
                         `4` = "20-25mm",
                         `5` = "25+mm"),
   eop_visibility_classification = recode(eop_visibility_classification,
                                          `0` = "eop_0mm",
                                          `1` = "0<_eop_<=5mm",
                                          `2` = "eop_>5mm"),
   fhp_category = recode(fhp_category,
                         `0` = "0-10mm",
                         `1` = "10-20mm",
                         `2` = "20-30mm",
                         `3` = "30-40mm",
                         `4` = "40-50mm",
                         `5` = "50-60mm",
                         `6` = "60-70mm",
                         `7` = "70-80mm")
   ) %>%
  select(sex, age, age_category, eop_size_mm, eop_category, eop_visibility_classification, eop_shape, eop_shape, fhp_size_mm, fhp_category, eeop)
```

* This dataset contains `r nrow(mtp_data)` observations and `r ncol(mtp_data)` variables including: sex; age and age category; eop category, size, classification, and shape; fhp size and category; and eeop.

### Issues: 
* Included a 17 y/o

```{r echo = FALSE}
mtp_data %>%
  filter(age == 17) %>%
  knitr::kable(caption = "Participant 7's Inclusion of 17-year-old Participant") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

* Several "eop_category" and "eop_visibility_classification" data were incorretly implemented (e.g., participant #3's "eop_size"='0', but "eop_visibility_classification"='>0' and #126's "eop_size_mm"='2.11' but "eop_size"='>5')

```{r echo = FALSE}
mtp_data %>%
  filter(age == 22, fhp_size_mm == 28.1) %>%
  knitr::kable(caption = "Participant 3's EOP Visibility Misclassification") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))

mtp_data %>%
  filter(eop_size_mm == 2.11) %>%
  knitr::kable(caption = "Participant 126's EOP Size Misclassification") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

* n=1221, but n=1200 reported in paper

```{r echo = FALSE}
mtp_data %>%
  count() %>%
  knitr::kable(caption = "Total Observations") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

* Poor variable definitions (e.g., eop_size: '1'='5-10', '2'='10-15')

### Age/Gender Distribution
```{r}
mtp_data %>%
  select(sex, age_category) %>%
  group_by(age_category, sex) %>%
  summarize(n = n()) %>%
  pivot_wider(
    names_from = sex,
    values_from = n
  ) %>%
  knitr::kable(caption = "Age Group and Gender Distribution") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

# Problem 2

```{r recreating the authors figure 3, echo = FALSE}
authors_figure_3 = 
  mtp_data %>%
  filter(fhp_size_mm > 0) %>%
  group_by(sex, age_category) %>%
  summarize(
    mean_fhp = mean(fhp_size_mm), 
    sd_fhp = sd(fhp_size_mm)
    ) %>%
  ggplot(aes(x = age_category, y = mean_fhp, fill = sex)) + 
  geom_errorbar(aes(x = age_category, ymin = (mean_fhp - sd_fhp), ymax = (mean_fhp + sd_fhp)), alpha = .5, width = .8, position = "dodge") +
  geom_histogram(stat = "identity", position = "dodge", binwidth = 1) +
  labs(
    title = "Recreation of the Author's Figure 3", 
    x = "Age Group (years)",
    y = "Forward Head Protraction Size (mm)", 
    caption = "Data from Scientific Reports Article by Shahar and Sayers") +
  scale_y_continuous(
    breaks = c(0, 10, 20, 30, 40, 50, 60),
    labels = c("0", "10", "20", "30", "40", "50", "60"))


authors_figure_3
```


```{r improved figure 3}
improved_figure_3 = 
  mtp_data %>%
  ggplot(aes(x = age_category, y = fhp_size_mm, color = sex)) + 
  geom_boxplot() +
  labs(
    title = "Forward Head Protraction Across \nAge Groups by Sex", 
    x = "Age Group (years)",
    y = "Forward Hear Protraction Size (mm)", 
    caption = "Data from Scientific Reports Article by Shahar and Sayers") +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80),
    labels = c("0", "20", "40", "60", "80"))

improved_figure_3
```

```{r}
improved_figure_3_another =
  mtp_data %>%
  ggplot(aes(x = age, y = fhp_size_mm, color = sex)) + 
  geom_point() +  
  geom_smooth(se = FALSE) +
  labs(
    title = "Association between Forward Head Protraction and Age by Sex", 
    x = "Age (years)",
    y = "Forward Hear Protraction Size (mm)", 
    caption = "Data from Scientific Reports Article by Shahar and Sayers") +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80),
    labels = c("0", "20", "40", "60", "80"))

improved_figure_3_another
```

### These figures show fhp_size doesn't really increase until age 55 and is only slightly larger in males.

```{r recreating the authors figure 4, echo = FALSE}
authors_figure_4 = 
  mtp_data %>%
  filter(eeop == "yes") %>%
  group_by(age_category, sex) %>%
  count() %>%
  ggplot(aes(x = age_category, y = n, group = sex)) + 
  geom_line() +
  geom_point(aes(shape = sex)) +
  labs(
    title = "Recreation of the Author's Figure 4", 
    x = "Age Group (years)",
    y = "EEOP Count", 
    caption = "Data from 2018 Scientific Reports Article by Shahar and Sayers") +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80, 100, 120),
    labels = c("0", "20", "40", "60", "80", "100", "120"))


authors_figure_4
```

```{r improved figure 4}
improved_figure_4 = 
  mtp_data %>%
  group_by(age_category, sex, eeop) %>%
  count() %>%
  pivot_wider(
    names_from = eeop,
    values_from = n
  ) %>%
  mutate(
    total = no + yes, 
    rate = (yes / total) * 100
  ) %>%
  ggplot(aes(x = age_category, y = rate, group = sex)) + 
  geom_line() +
  geom_point(aes(shape = sex)) +
  labs(
    title = "Rate of Enlarged External Occipital \nProturberance for each Age Group and Sex", 
    x = "Age Group (years)",
    y = "EEOP Rate (percent)", 
    caption = "Data from 2018 Scientific Reports Article by Shahar and Sayers") +
  scale_y_continuous(
    breaks = c(0, 10, 20, 30, 40, 50, 60, 70),
    labels = c("0", "10%", "20%", "30%", "40%", "50%", "60%", "70%"))

improved_figure_4
```

### The eeop rate is higher in every age group for males, and is highest among 18-30 y/o's.

## Two-Paneled Figure

```{r}
improved_figure_3 + improved_figure_4
```

## FHP/EOP Association

```{r}
fhp_eop_association = 
  mtp_data %>%
  filter(
    eop_size_mm > 0
    ) %>%
  ggplot(aes(x = eop_size_mm, y = fhp_size_mm, color = sex)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(~age_category) + 
  labs(
    title = "Association between Forward Head Protraction and \nExternal Occipital Proturberance Size Across Age Groups by Sex", 
    x = "External Occipital Proturberance (mm)",
    y = "Forward Head Protraction (mm)", 
    caption = "Data from 2018 Scientific Reports Article by Shahar and Sayers")
fhp_eop_association
```

### There's a positive association between FHP and EOP in each age group for males, but only in the 18-30 and 61+ for females. The association is strongest in those 61+.

### "eop_size" = '0' was removed as these are missing values. Including '0's would falsely make all associations positive: 

```{r echo = FALSE}
fhp_eop_association = 
  mtp_data %>%
  ggplot(aes(x = eop_size_mm, y = fhp_size_mm, color = sex)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(~age_category) + 
  labs(
    title = "Association between Forward Head Protraction and \nExternal Occipital Proturberance Size Across Age Groups by Sex", 
    x = "External Occipital Proturberance (mm)",
    y = "Forward Head Protraction (mm)", 
    caption = "Data from 2018 Scientific Reports Article by Shahar and Sayers")
fhp_eop_association
```

# Problem 3

## Sample Size

```{r}
mtp_data %>%
  select(age_category) %>%
  group_by(age_category) %>%
  summarize(n = n()) %>%
  knitr::kable(caption = "Age Distribution of Sample") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

### Total n (1221 vs. 1200) and age group sizes were reported differently than our results (e.g., 18-30: 303 vs 300).

## FHP

```{r}
mtp_data %>%
  filter(fhp_size_mm > 0) %>%
  group_by(sex) %>%
  summarize(
    mean_fhp = mean(fhp_size_mm), 
    sd_fhp = sd(fhp_size_mm)
    ) %>%
  knitr::kable(digits = 1, caption = "Mean and Standard Deviation of Sample by Gender") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))

mtp_data %>%
  filter(fhp_size_mm > 0) %>%
  summarize(
    mean_fhp = mean(fhp_size_mm), 
    sd_fhp = sd(fhp_size_mm)
    ) %>%
  knitr::kable(digits = 1, caption = "Mean and Standard Deviation of Entire Sample") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

### The reported mean FHP by gender and overall was slightly different than our results.

## EEOP Prevalence

```{r}
mtp_data %>%
  group_by(eeop) %>%
  count() %>%
  pivot_wider(
    names_from = eeop,
    values_from = n
  ) %>%
  mutate(
    total = no + yes, 
    rate = (yes / total) * 100
  ) %>%
  knitr::kable(digits = 1, caption = "Prevalence of EEOP") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", font_size = 12))
```

### The definition of EEOP is "bony outgrowths >10mm" but "eop_size" and "eop_visibility_classification" were not categorized this way. 
### A new variable "eeop" was created during data cleaning to meet this definition. 
### The 32.3% prevalence is close to the reported 33%. 

## FHP Trends

```{r}
mtp_data %>%
  mutate(
    fhp_40 = recode(fhp_category,
                    `0-10mm` = "no",
                    `10-20mm` = "no",
                    `20-30mm` = "no",
                    `30-40mm` = "no",
                    `40-50mm` = "yes",
                    `50-60mm` = "yes",
                    `60-70mm` = "yes",
                    `70-80mm` = "yes",)
  ) %>%
  group_by(fhp_40, age_category) %>%
  count() %>%
  pivot_wider(
    names_from = fhp_40,
    values_from = n
  ) %>%
  mutate(
    total = no + yes, 
    rate = (yes / total) * 100
  ) %>%
  ggplot(aes(x = age_category, y = rate)) +
  geom_histogram(stat = "identity") +
  labs(
    title = "Rate of Pronounced Forward Head Protraction (FHP, >40mm) \nAcross Age Groups", 
    x = "Age Group (years)",
    y = "Frequency of FHP >40mm", 
    caption = "Data from Scientific Reports Article by Shahar and Sayers")
```

### Again, the numbers are slightly different (32.5% vs 34.5% reported). 
### Their claim is consistent with our data (FHP>40 mm was observed frequently in those 61+)
### Broadly speaking, there's an increase in FHP for those 61+.

# Problem 4

### FHP increases with age, is higher in males (as is eeop rate), and positively associated with EOP size in those 61+; however, the authors contort the data to make their results seem more extreme. 
### They excluded participants without explanation and miscoded categorical variables which creates doubts about the quality of their analysis.
### It's impossible to know if cellphones cause horn growth from this poorly analyzed retrospective study (causation only possible in RCT's). 